---
title: "Accessible Data Analysis: Data Exploration"
author: "Christy Garcia and Chris Prener"
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This is the second of three notebooks for our ATLAS week session.

## Dependencies
The following packages are required for this notebook:

```{r load-dependencies}
# tidyverse packages
library(dplyr)  # data wrangling
library(readr)  # data import

# other packages
library(here)    # file paths
library(janitor) # frequency tables
library(skimr)   # descriptive statistics
```

## Loading Data
We need the same two data tables we used in the first notebook for this notebook as well:

```{r}
lead <- read_csv(here("data", "stlLead.csv"))
murders <- read_csv(here("data", "stlMurders.csv"))
```

## Exploring Data
### Review
In the first notebook, we used the `glimpse()` function from `dplyr` to get a list of all variables present in our data frame:

```{r}
glimpse(murders)
```

This is an easy way to get some basic information about each variable.

### Descriptive Statistics
The `skimr` package provides some numeric summaries of the data present in a given data frame. The `skim()` function can be used to generate a report that breaks down variables by type, and provides the most relevant descriptive statistics based on that distinction:

```{r}
skim(murders)
```

Now you try generating some descriptive statistics for the `lead` data:

```{r}
skim(lead)
```

### Pipes
All of the `tidyverse` packages and other associated packages, like `janitor` (introduced below), use the pipe operator `%>%`. The advantage of using the pipe is that it makes the code easier to read by "piping" a series of functions together in a way that can be read by a paragraph:

1. We take the `murders` data, **then**
2. we create a new variable named `northSide`, **then**
3. we create a new variable named `greaterVille`, and
4. assign the results back onto the `murders` data frame.

```{r}
murders %>%
  mutate(northSide = ifelse(nhood > 45, "North City", "South City")) %>%
  mutate(greaterVille = ifelse(nhood >= 50 & nhood <= 59, TRUE, FALSE)) -> murders
```

Pipes don't necessarily make our code more efficient, but they make it more human readable.

### Frequency Tables
The `janitor` package makes it easy to create frequency tables for string and factor data. We use the `tabyl()` function to accomplish this. At its simplest, `tably()` takes just the variable name when included at the end of a pipe:

```{r}
murders %>%
  tabyl(year)
```

We can add "adornments", including a total row at the bottom and well-formatted percent symbols:

```{r}
murders %>%
  tabyl(year) %>%
  adorn_totals() %>%
  adorn_pct_formatting()
```

Now its your turn. Create a frequency table of the `northSide` variable using pipes and `janitor`:

```{r}
murders %>%
  tabyl(northSide) %>%
  adorn_totals() %>%
  adorn_pct_formatting()
```

