---
title: "Atlas Week 2018"
subtitle: "An Introduction to `R`, RStudio, and the `tidyverse`"
author: "SLU DSS"
date: "04/11/2018"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

class: inverse, center, middle

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

# Overview

---
# What is the SLU DSS?

The [SLU Data Science Seminar](https://slu-dss.githb.io) (DSS) is a collaborative, interdisciplinary group at Saint Louis University focused on building researchers’ data science skills using open source software. We currently host seminars focused on the programming language `R`.

We host:

1. Four-part seminars each semester on specific topics related to `R`
2. Brownbags by guest facilitators showcasing how they use `R` for research or teaching
3. An ATLAS Week Session introducing participants to `R`

---
# Basics

- **`R`** is a programming language first developed in 1993 for statistical computing
    - The `R` ecosystem has rapidly expanded in the last ten years to include tools for project management, graphics production, machine learning, GIS, and dissemination (these slides are made with `R`!)
- People are attracted to `R` because it is:
    - *free*
    - *highly extensible* - there are over 12,000 packages available on **CRAN** 

--
- ...but `R` has a bit of a reputation...
    
--

- RStudio is a **G**raphical **U**ser **I**nterface (**GUI**) for `R` that makes working with the programming language *much* easier

--

- One of the reasons interest in `R` has grown rapidly is the **`tidyverse`**, which is a family of packages that use an easy to read and implement syntax for conducting common data management tasks

---
# Installing Packages

One of the first tasks we face with `R` is getting packages installed. The `base` distribution of `R` comes with some packages, but we'll need a few more for today. Each package contains a set of *functions* (easy to execute, generalized blocks of code) that we can use in our own projects.

The `install.packages()` function from `utils` is the easiest way to install a package:

```{r install-packages, eval=FALSE}
install.packages("usethis")
```

--

We only need to install packages once per user, but we may need to update them periodically:

```{r update-packages, eval=FALSE}
update.packages()
```

---
# Loading Packages

While we only have to install packages once per user, we must load them at the beginning of **each** session using the `library()` function:

```{r load-packages}
library(usethis)
```
